### 2019-03-06 2019_csic5011/slides/homework03.tex
Add files via upload

\documentclass[11pt]{article}

\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}

\def\N{{\mathbb N}}
\def\NN{{\mathcal N}}
\def\R{{\mathbb R}}
\def\E{{\mathbb E}}
\def\tr{{\mathrm{trace}}}
\def\P{{\mathrm{Prob}}}
\def\sign{{\mathrm{sign}}}

\setlength{\oddsidemargin}{0.25 in}
\setlength{\evensidemargin}{-0.25 in}
\setlength{\topmargin}{-0.6 in}
\setlength{\textwidth}{6.5 in}
\setlength{\textheight}{8.5 in}
\setlength{\headsep}{0.75 in}
\setlength{\parindent}{0.25 in}
\setlength{\parskip}{0.1 in}

\newcommand{\lecture}[4]{
   \pagestyle{myheadings}
   \thispagestyle{plain}
   \newpage
   \setcounter{page}{1}
   \setcounter{section}{0}
   \noindent
   \begin{center}
   \framebox{
      \vbox{\vspace{2mm}
    \hbox to 6.28in { {\bf A Mathematical Introduction to Data Science \hfill #4} }
       \vspace{6mm}
       \hbox to 6.28in { {\Large \hfill #1  \hfill} }
       \vspace{6mm}
       \hbox to 6.28in { {\it Instructor: #2\hfill #3} }
      \vspace{2mm}}
   }
   \end{center}
   \markboth{#1}{#1}
   \vspace*{4mm}
}

\begin{document}

\lecture{Homework 3. High Dimensional Statistics Models}{Yuan Yao}{Due: Open Date}{March 6, 2019}

The problem below marked by $^*$ is optional with bonus credits. For the experimental problem, include the source codes which are runnable under standard settings. Since there is NO grader assigned for this class, homework will not be graded. But if you would like to submit your exercise, please send your homework to the address (\href{mailto:datascience.hw@gmail.com}{datascience.hw@gmail.com})  with a title ``CSIC5011: Homework \#". I'll read them and give you bonus credits.

\begin{enumerate}
\item {\em Maximum Likelihood Method}: consider $n$ random samples from a multivariate normal distribution, $X_i \in \R^p \sim \NN(\mu,\Sigma)$ with $i=1,\ldots,n$.

\begin{enumerate}
\item Show the log-likelihood function
\[ l_n(\mu,\Sigma) = - \frac{n}{2} \tr(\Sigma^{-1} S_n ) - \frac{n}{2} \log \det(\Sigma) + C, \]
where $S_n = \frac{1}{n}\sum_{i=1}^n (X_i - \mu)(X_i - \mu)^T$, and some constant $C$ does not depend on $\mu$ and $\Sigma$; 
\item Show that $f(X) = \tr(AX^{-1})$ with $A, X\succeq 0$ has a first-order approximation,
\[ f(X+\Delta) \approx f(X) - \tr(X^{-1}A' X^{-1} \Delta ) \]
hence formally $d f(X) / d X = - X^{-1} A X^{-1}$ (note $(I+X)^{-1} \approx I - X$. A typo in previous version missed `-' sign here.);  
\item Show that $g(X) = \log\det(X)$ with $A, X\succeq 0$ has a first-order approximation,
\[ g(X + \Delta) \approx g(X) + \tr(X^{-1} \Delta) \]
hence $d g(X) / d X = X^{-1}$ (note: consider eigenvalues of $X^{-1/2} \Delta X^{-1/2}$); 
\item Use these formal derivatives with respect to positive semi-definite matrix variables to show that the maximum likelihood estimator of $\Sigma$ is 
\[ \hat{\Sigma}^{MLE}_n = S_n. \]
\end{enumerate}
A reference for (b) and (c) can be found in Convex Optimization, by Boyd and Vandenbergh, examples in Appendix A.4.1 and A.4.3:

\url{https://web.stanford.edu/~boyd/cvxbook/bv_cvxbook.pdf}

\item {\em Shrinkage:} Suppose $y\sim \NN(\mu, I_p)$.

\begin{enumerate}
\item Consider the Ridge regression
\[ \min_{\mu} \frac{1}{2}\|y - \mu \|_2^2 +\frac{\lambda}{2} \|\mu\|_2^2. \]
Show that the solution is given by 
\[ \hat{\mu}^{ridge}_i = \frac{1}{1+\lambda} y_i .\]
Compute the risk (mean square error) of this estimator. The risk of MLE is given when $C=I$. 
\item Consider the LASSO problem,
\[ \min_{\mu} \frac{1}{2}\|y - \mu \|_2^2 +\lambda \|\mu\|_1. \]
Show that the solution is given by Soft-Thresholding
\[ \hat{\mu}^{soft}_i = \mu_{soft}( y_i;\lambda) := \sign( y_i ) (|y_i |- \lambda)_+. \]
For the choice $\lambda = \sqrt{2\log p}$, show that the risk is bounded by
\[ \E\|\hat{\mu}^{soft}(y) -\mu\|^2  \leq 1+ (2\log p + 1) \sum_{i=1}^p \min(\mu_i^2,1).\]
Under what conditions on $\mu$, such a risk is smaller than that of MLE?
Note: see Gaussian Estimation by Iain Johnstone, Lemma 2.9 and the reasoning before it.
\item Consider the $l_0$ regularization 
\[ \min_{\mu} \|y - \mu \|_2^2 +\lambda^2 \|\mu\|_0, \]
where $\|\mu\|_0 := \sum_{i=1}^p I(\mu_i \neq 0)$. Show that the solution is given by Hard-Thresholding
\[ \hat{\mu}^{hard}_i = \mu_{hard}( y_i;\lambda) := y_i  I(|y_i|>\lambda).\]
Rewriting $\hat{\mu}^{hard}(y) = (1-g(y))y$, is $g(y)$ weakly differentiable? Why? 
\item Consider the James-Stein Estimator 
\[ \hat{\mu}^{JS}(y) = \left( 1 - \frac{\alpha}{\|y\|^2} \right) y. \]
Show that the risk is 
\[ \E\|\hat{\mu}^{JS}(y) - \mu\|^2 = \E U_\alpha(y) \]
where $U_\alpha(y) = p - (2\alpha(p-2) - \alpha^2) / \|y\|^2$. Find the optimal $\alpha^\ast=\arg\min_\alpha U_\alpha(y)$. Show that for $p>2$, the risk of James-Stein Estimator is smaller than that of MLE for all $\mu\in \R^p$. 
\item In general, an odd monotone unbounded function $\Theta:\R\to\R$ defined by $\Theta_\lambda(t)$ with parameter $\lambda \geq 0$ is called \emph{shrinkage} rule, if it satisfies 
\subitem[shrinkage] $0\leq \Theta_\lambda(|t|) \leq |t|$; 
\subitem[odd] $\Theta_\lambda(-t) = - \Theta_\lambda(t)$;
\subitem[monotone] $\Theta_\lambda(t) \leq \Theta_\lambda(t^\prime)$ for $t\leq t^\prime$;
\subitem[unbounded] $\lim_{t\to \infty} \Theta_\lambda(t)=\infty$. \\
Which rules above are shrinkage rules? 
\end{enumerate}

\item {\em *Necessary Condition for Admissibility of Linear Estimators}. Consider linear estimator for $y\sim \NN(\mu,\sigma^2 I_p)$
\[ \hat{\mu}_C(y) = C y. \]
Show that $\hat{\mu}_C$ is admissible only if
		\begin{enumerate}
			\item $C$ is symmetric;
			\item $0 \le \rho_i(C) \le 1$ (where $\rho_i(C)$ are eigenvalues of $C$);
			\item $\rho_i(C) = 1$ for at most two $i$.
		\end{enumerate}	
		These conditions are satisfied for MLE estimator when $p=1$ and $p=2$.

Reference: Theorem 2.3 in Gaussian Estimation by Iain Johnstone,\\
\url{http://statweb.stanford.edu/~imj/Book100611.pdf}

\item {\em James Stein Estimator for $p=1$:}

From Theorem 3.1 in the lecture notes, we know that MLE $\hat{\mu} = Y$ is admissible when $p=1 \text{ or } 2$. However if we use SURE to calculate the risk of James Stein Estimator,
\[
	R(\hat{\mu}^{\text{JS}},\mu) = \E U(Y) 
	= p - \E_\mu \frac{(p-2)^2}{\lVert Y \rVert^2}
	< p = R(\hat{\mu}^{\text{MLE}},\mu)
\]
it seems that for $p=1$ James Stein Estimator should still has lower risk than MLE for any $\mu$.
 Explain what violates the above calculation for $p=1$.

\item {\em Phase transition in PCA ``spike" model:} Consider a finite sample of $n$ i.i.d vectors $x_1,x_2,\ldots,x_n$ drawn from the $p$-dimensional Gaussian distribution $\mathcal{N}(0,\sigma^2 I_{p\times p} + \lambda_0 uu^T)$, where $\lambda_0/\sigma^2$ is the signal-to-noise ratio (SNR) and $u \in \mathbb{R}^p$. In class we showed that the largest eigenvalue $\lambda$ of the sample covariance matrix $S_n$
    $$S_n = \frac{1}{n} \sum_{i=1}^n x_i x_i^T$$
    pops outside the support of the Marcenko-Pastur distribution if $$\frac{\lambda_0}{\sigma^2} > \sqrt{\gamma},$$ or equivalently, if $$\text{SNR} > \sqrt{\frac{p}{n}}.$$ (Notice that $\sqrt{\gamma} < (1+\sqrt{\gamma})^2$, that is, $\lambda_0$ can be ``buried" well inside the support Marcenko-Pastur distribution and still the largest eigenvalue pops outside its support). All the following questions refer to the limit $n\to \infty$ and to almost surely values:

\begin{enumerate}
\item Find $\lambda$ given $\text{SNR} > \sqrt{\gamma}$.
\item Use your previous answer to explain how the SNR can be estimated from the eigenvalues of the sample covariance matrix.
\item Find the squared correlation between the eigenvector $v$ of the sample covariance matrix (corresponding to the largest eigenvalue $\lambda$) and the ``true" signal component $u$, as a function of the SNR, $p$ and $n$. That is, find $|\langle u,v \rangle|^2$. 
\item Confirm your result using MATLAB or R simulations (e.g. set $u = e$; and choose $\sigma=1$ and $\lambda_0$ in different levels. Compute the largest eigenvalue and its associated eigenvector, with a comparison to the true ones.)
\end{enumerate}

\item {\em Exploring S$\&$P500 Stock Prices:} Take the Standard \& Poor's 500 data: \\
\url{http://math.stanford.edu/~yuany/course/data/snp452-data.mat},\\
which contains the data matrix $X\in \R^{n\times p}$ of $n=1258$ consecutive observation days and $p=452$ daily closing stock prices, and the cell variable ``${\rm stock}$" collects the names, codes, and the affiliated industrial sectors of the 452 stocks. Use Matlab or R for the following exploration. 
\begin{enumerate}
\item Take the logarithmic prices $Y=\log X$; 
\item For each observation time $t\in \{1,\ldots,1257\}$, calculate logarithmic price jumps
\[ \Delta Y_{i,t} = Y_{i,t} - Y_{i,{t-1}}, \ \ \ i\in \{ 1,\ldots, 452 \}; \]
\item Construct the realized covariance matrix $\hat{\Sigma}\in\R^{452\times 452}$ by,
\[ \hat{\Sigma}_{i,j} = \frac{1}{1257} \sum_{\tau =1}^{1257} \Delta Y_{i,\tau} \Delta Y_{j, \tau}; \]
\item Compute the eigenvalues (and eigenvectors) of $\hat{\Sigma}$ and store them in a descending order by $\{\hat{\lambda}_k, k=1,\ldots,p\}$. 
\item \emph{Horn's Parallel Analysis}: the following procedure describes a so-called Parallel Analysis of PCA using random permutations on data. Given the matrix $[\Delta Y_{i,t}]$, apply random permutations $\pi_i: \{1,\ldots,t\} \to \{1,\ldots, t\}$ on each of its rows: $\Delta \tilde{Y}_{i,\pi_i(j)}$ such that 
\[
[\Delta \tilde{Y}_{\pi(i),t}] = \left[
\begin{array}{ccccc}
\Delta Y_{1,1} & \Delta Y_{1,2} & \Delta Y_{1,3} & \ldots & \Delta Y_{1,t} \\
\Delta Y_{2,\pi_2(1)} & \Delta Y_{2,\pi_2(2)} & \Delta Y_{2,\pi_2(3)} & \ldots & \Delta Y_{2,\pi_2(t)} \\
\Delta Y_{3,\pi_3(1)} & \Delta Y_{3,\pi_3(2)} & \Delta Y_{3,\pi_3(3)} & \ldots & \Delta Y_{3,\pi_3(t)} \\
\ldots & \ldots & \ldots & \ldots & \ldots \\
\Delta Y_{n,\pi_n(1)} & \Delta Y_{n,\pi_n(2)} & \Delta Y_{n,\pi_n(3)} & \ldots & \Delta Y_{n,\pi_n(t)} 
\end{array}
\right].
\]
Define $\tilde{\Sigma} = \frac{1}{t} \Delta \tilde{Y} \cdot \Delta \tilde{Y}^T$ as the null covariance matrix. Repeat this for $R$ times and compute the eigenvalues of $\tilde{\Sigma}_r$ for each $1\leq r\leq R$. Evaluate the $p$-value for each estimated eigenvalue $\hat{\lambda}_k$ by $(N_k + 1)/(R+1)$ where $N_k$ is the counts that $\hat{\lambda}_k$ is less than the $k$-th largest eigenvalue of $\tilde{\Sigma}_r$ over $1\leq r \leq R$. Eigenvalues with small $p$-values indicate that they are less likely arising from the spectrum of a randomly permuted matrix and thus considered to be signal. Draw your own conclusion with your observations and analysis on this data. A reference is: Buja and Eyuboglu, "Remarks on Parallel Analysis", Multivariate Behavioral Research, 27(4): 509-540, 1992.
\end{enumerate}

\item *{\em Finite rank perturbations of random symmetric matrices:} Wigner's semi-circle law (proved by Eugene Wigner in 1951) concerns the limiting distribution of the eigenvalues of random symmetric matrices. It states, for example, that the limiting eigenvalue distribution of $n\times n$ symmetric matrices whose entries $w_{ij}$ on and above the diagonal $(i\leq j)$ are i.i.d Gaussians $\mathcal{N}(0,\frac{1}{4n})$ (and the entries below the diagonal are determined by symmetrization, i.e., $w_{ji}=w_{ij}$) is the semi-circle:
    $$p(t) = \frac{2}{\pi} \sqrt{1-t^2}, \quad -1\leq t \leq 1,$$
    where the distribution is supported in the interval $[-1,1]$.
\begin{enumerate}
\item Confirm Wigner's semi-circle law using MATLAB or R simulations (take, e.g., $n=400$).
\item Find the largest eigenvalue of a rank-1 perturbation of a Wigner matrix. That is, find the largest eigenvalue of the matrix $$W + \lambda_0 uu^T,$$ where $W$ is an $n\times n$ random symmetric matrix as above, and $u$ is some deterministic unit-norm vector. Determine the value of $\lambda_0$ for which a phase transition occurs. What is the correlation between the top eigenvector of $W+\lambda_0 uu^T$ and the vector $u$ as a function of $\lambda_0$? Use techniques similar to the ones we used in class for analyzing finite rank perturbations of sample covariance matrices.
\end{enumerate}

[Some Hints about homework] For Wigner Matrix\ $W=[w_{ij}]_{n\times n},w_{ij}=w_{ji},w_{ij}\sim N(0,\frac{\sigma}{\sqrt{n}})$, 
 the answer is
$$
\begin{array}{rcl}
\textrm{eigenvalue\ is}\ & &\lambda=R+\frac{1}{R}\\
\textrm{eigenvector\ satisfies}\ && (u^{T}\hat{v})^{2}=1-\frac{1}{R^{2}} \\
\end{array}
$$

\end{enumerate}

\end{document}