### 2023-04-16 course/csic5011/2023/project1/Hao-Hu-Li-Liu/Code1.Rmd
project1

---
title: "Reduction, visualization and classification"
author: "Liu Yonglin"
date: "2023-04-09"
output: html_document
---

```{r setup, include=FALSE}
library(R.matlab)
library(RDRToolbox)
library(ggplot2)
library(ggcorrplot)
library(Rtsne)
#library(rpca)
library(rospca)
library(sparsepca)
#library(lle)
library(randomForest)
library(pROC)
library(nnet)
```

## set seed
```{r}
set.seed(42)
```

## Load data
```{r}
path<-getwd()
path<-paste(path,"snp452-data.mat",sep = "/")
datmat<-readMat(path)
#dat<-load("/snp500.Rda")

x<-as.data.frame(datmat$X)
y<-as.data.frame(t(array(unlist(datmat$stock),dim = c(3,452))))#name/full name/group
colnames(y)<-c("name","detail","group")
colnames(x)<-y[,1]

```

# Distribution of stock class
```{r}
library(ggplot2)
z<-unique(y$group)
n<-length(z)
gp<-rep(0,n)
for(i in 1:n){
  gp[i]<-length(which(y$group==z[i]))
}

count_bar<-data.frame("group"=z,"frequency"=gp)
ggplot(data = count_bar,aes(x=factor(group),y=frequency,fill=factor(group)))+geom_bar(stat = "identity")+xlab("Groups")+ylab("Frequency")+theme(axis.text.x = element_blank(),legend.key.size=unit(10,"pt"))+guides(fill=guide_legend(title = "Group"))
```

## Class
```{r}
target<-rep(0,10)
index_all<-matrix(rep(0,10*100),nrow = 10)

for(i in 1:10){
  index<-count_bar$group[i]
  print(index)
  print(which(y$group==index)) ###
  ii=as.vector(which(y$group==index))
  index_all[i,]=c(ii,rep(0,100-gp[i]))
  target[i]<-which(y$group==index)[1]
}
```

## Select the first company in each class
```{r}
target_gp<-count_bar$group
```

## Price of stocks by days
```{r}
days=as.data.frame(c(1:1258))
price <- data.frame(matrix(ncol = 1258, nrow = 0))
dd <- data.frame(matrix(ncol = 1258, nrow = 0))
for(i in 1:10){
  price<-rbind(price,data.frame(x[,target[i]], "group"=target_gp[i]))
  dd=rbind(dd,days)
  #plot(1:1258,as.numeric(x[,target[i]]),type = "l",main = count_bar$group[i],xlab="Days",ylab="Stock Prices")
}
price=cbind(dd,price)
colnames(price)<-c("x","y","group")
ggplot(data = price,aes(x=x, y = y, col=factor(group)))  +geom_line()

#ggplot(data=new.data,aes(x=name,y=mean_value))+
#  geom_line(aes(color=variants,lty=`Reference genome`))+
#  geom_point(aes(color=variants))

```

# Growth rate of selected companies
```{r}
rate<-matrix(rep(0,10*1257),nrow = 10)

for(i in 1:10){

  x1=x[1:1257,target[i]]
  x2=x[2:1258,target[i]]
  dx=(x2-x1)/x1

  rate[i,]=dx

  plot(1:1257,as.numeric(dx),type = "l",main = count_bar$group[i],xlab="Days",ylab="Growth Rates")

  hist(scale(dx),prob=T)

  xfit<-seq(min(scale(dx)),max(scale(dx)),length=40)
  yfit<-dnorm(xfit,mean(scale(dx)),sd(scale(dx)))
  lines(xfit,yfit,col="red",lwd=3)
}
```

# Growth rate of all companies
```{r}
index_all=t(index_all)

rate<-matrix(rep(0,452*1257),nrow = 452)

for(i in 1:10){ #i:种类 j:每个种类中的企业
  for(j in 1:gp[i]){
  x1=x[1:1257,index_all[j,i]]
  x2=x[2:1258,index_all[j,i]]
  dx=(x2-x1)/x1

  rate[i,]=dx

  rate[index_all[j,i],]=as.vector(dx)
  }
}

#
q=matrix(rep(0,1*452),nrow = 1)
for (i in 1:10){
  for(j in 1:gp[i]){
    q[index_all[j,i]]=z[i]
  }
}
q=t(q)

#ratelle=rate
```

## Plot heatmaps
```{r}
## Heatmap
#corr_matrix <- cor(X3)
#ggcorrplot(corr_matrix)
```

## Heatmap in groups by Yifan
```{r}
#heatmap in-group
library(ggcorrplot)
target<-target[-9]
target_gp<-count_bar$group[-9]
h<-matrix(rep(0,452*452),ncol = 452)
for(i in 1:452){
  for(j in i:452){
    h[i,j]<-cor(x[,i],x[,j])
    h[j,i]<-h[i,j]
  }
}

for(i in 1:9){
  index<-which(y$group==target_gp[i])
  plt<-ggcorrplot(h[index,index])
  print(plt)
}
```

### ISOMAP

##Create data frame with 0 rows and 3 columns
```{r}
PC <- data.frame(matrix(ncol = 3, nrow = 0))
```

## Data reduction and creating new dataframe PC
```{r}
for (i in 1:10){

  index<-which(y$group==target_gp[i])

  new=rate[index,]
  #new=t(new)
  new=as.matrix(new)

  X_ISO<-Isomap(data=new, dims=2, k=4)
  X_ISO<-as.data.frame(X_ISO)

  PC<-rbind(PC,data.frame("x"=X_ISO$dim2.1, "y"=X_ISO$dim2.2, "group"<-rep(i,gp[i])))
}

```

## provide column names
```{r}
colnames(PC)<-c("PC1","PC2","group")
```

## splitting
```{r}
set.seed(42)
sample=sample(c(TRUE, FALSE), nrow(PC), replace=TRUE, prob=c(0.7,0.3))
isotrain  <- PC[sample, ]
isotest   <- PC[!sample, ]
xtrain_iso=isotrain[,1:2]
xtest_iso=isotest[,1:2]
ytrain_iso=isotrain[,3]
ytest_iso=isotest[,3]
```

#logistic regression
```{r}
lr_iso<-multinom(group~PC1+PC2,data=isotrain,maxit=500) #maxit:iteration times
summary(lr_iso) 
```

## Error and accuracy
```{r}
prob_isotrain<-predict(lr_iso,type="probs") 
head(prob_isotrain)

pred_isotrain<-predict(lr_iso,type="class") 
head(pred_isotrain)

#Confusion Matrix
table<-table(predicted=pred_isotrain,Actual=isotrain$group) 
table

#Accuracy
Accuracy<-sum(diag(table))/sum(table)  
Accuracy

#lr_iso=glm(group~PC1+PC2,data=isotrain)

#summary(lr_iso)

```

## Plot
```{r}
ggplot(data = PC,aes(x=PC1, y = PC2, col=factor(group)))+geom_point()
```

## Save
```{r}
v1=PC[1:452,1]
v2=PC[1:452,2]
d2=matrix(rep(0,452*452),nrow = 452)
for (i in 1:452){
  for (j in 1:452){
    d2[i,j]=(v1[i]-v1[j])^2+(v2[i]-v2[j])^2
  }
}

write.csv(d2,"iso3.csv")
```

### LLE

##Create data frame with 0 rows and 3 columns
```{r}
PC <- data.frame(matrix(ncol = 3, nrow = 0))
```

## Data reduction
```{r}
for (i in 1:10){

  index<-which(y$group==target_gp[i])

  new=rate[index,]
  #new=t(new)
  new=as.matrix(new)

  X_LLE<-LLE(data=new, dim=2, k=2)
  X_LLE<-as.data.frame(X_LLE)

  PC<-rbind(PC,data.frame("x"=X_LLE$V1, "y"=X_LLE$V2, "group"<-rep(target_gp[i],gp[i])))
}
```

##provide column names
```{r}
colnames(PC)<-c("PC1","PC2","group")
```

##Plot
```{r}
ggplot(data = PC,aes(x=PC1, y = PC2, col=factor(group)))+geom_point()
```

## splitting
```{r}
set.seed(42)
sample=sample(c(TRUE, FALSE), nrow(PC), replace=TRUE, prob=c(0.7,0.3))
isotrain  <- PC[sample, ]
isotest   <- PC[!sample, ]
xtrain_iso=isotrain[,1:2]
xtest_iso=isotest[,1:2]
ytrain_iso=isotrain[,3]
ytest_iso=isotest[,3]
```

#logistic regression
```{r}
lr_iso<-multinom(group~PC1+PC2,data=isotrain,maxit=500) #maxit:iteration times
summary(lr_iso) 
```

## Error and accuracy
```{r}
prob_isotrain<-predict(lr_iso,type="probs") 
head(prob_isotrain)

pred_isotrain<-predict(lr_iso,type="class") 
head(pred_isotrain)

#Confusion Matrix
table<-table(predicted=pred_isotrain,Actual=isotrain$group) 
table

#Accuracy
Accuracy<-sum(diag(table))/sum(table)  
Accuracy

#lr_iso=glm(group~PC1+PC2,data=isotrain)

#summary(lr_iso)

```

## Save data to csv
```{r}
v1=PC[1:452,1]
v2=PC[1:452,2]
d2=matrix(rep(0,452*452),nrow = 452)
for (i in 1:452){
  for (j in 1:452){
    d2[i,j]=(v1[i]-v2[j])^2
  }
}

write.csv(d2,"LLE2.csv")
```

### MDS

## create data frame with 0 rows and 3 columns
```{r}
PC <- data.frame(matrix(ncol = 3, nrow = 0))
```

## Data reduction
```{r}
for (i in 1:10){

  index<-which(y$group==target_gp[i])

  new=rate[index,]
  #new=t(new)

  dist = dist(new,p = 2)
  X_MDS= cmdscale(dist)
  X_MDS<-as.data.frame(X_MDS)

  PC<-rbind(PC,data.frame("x"=X_MDS$V1, "y"=X_MDS$V2, "group"<-rep(i,gp[i])))

}
```

## provide column names
```{r}
colnames(PC)<-c("PC1","PC2","group")
```

## Plot
```{r}
ggplot(data = PC,aes(x=PC1, y = PC2, col=factor(group)))+geom_point()
```

## splitting
```{r}
set.seed(42)
sample=sample(c(TRUE, FALSE), nrow(PC), replace=TRUE, prob=c(0.7,0.3))
mdstrain  <- PC[sample, ]
mdstest   <- PC[!sample, ]
xtrain_mds=mdstrain[,1:2]
xtest_mds=mdstest[,1:2]
ytrain_mds=mdstrain[,3]
ytest_mds=mdstest[,3]
```

## logistic regression
```{r}
lr_mds<-multinom(group~PC1+PC2,data=mdstrain,maxit=500) #maxit表示迭代次数
summary(lr_mds) 
```

## Accuracy
```{r}
prob_mdstrain<-predict(lr_mds,type="probs")   
head(prob_mdstrain)

pred_mdstrain<-predict(lr_mds,type="class")   
head(pred_mdstrain)
table<-table(predicted=pred_mdstrain,Actual=mdstrain$group)
table
Accuracy<-sum(diag(table))/sum(table)
Accuracy

```

## Save
```{r}
v1=PC[1:452,1]
v2=PC[1:452,2]
d2=matrix(rep(0,452*452),nrow = 452)
for (i in 1:452){
  for (j in 1:452){
    d2[i,j]=(v1[i]-v1[j])^2+(v2[i]-v2[j])^2
  }
}

write.csv(d2,"mds3.csv")
```

### TSNE

## create data frame with 0 rows and 3 columns
```{r}
PC <- data.frame(matrix(ncol = 3, nrow = 0))
```

#reduction
```{r}
for (i in 1:10){

  index<-which(y$group==target_gp[i])

  new=rate[index,]
  #new=t(new)
  new=as.matrix(new)

  tSNE_out = Rtsne(new,check_duplicates = FALSE, perplexity = 1)
  X_tSNE = tSNE_out$Y
  X_tSNE<-as.data.frame(X_tSNE)

  PC<-rbind(PC,data.frame("x"=X_tSNE$V1, "y"=X_tSNE$V2, "group"<-rep(i,gp[i])))

}
```

##provide column names
```{r}
colnames(PC)<-c("PC1","PC2","group")
```

##Plot
```{r}
ggplot(data = PC,aes(x=PC1, y = PC2, col=factor(group)))+geom_point()
```

## splitting
```{r}
set.seed(42)
sample=sample(c(TRUE, FALSE), nrow(PC), replace=TRUE, prob=c(0.7,0.3))
tsnetrain  <- PC[sample, ]
tsnetest   <- PC[!sample, ]
xtrain_tsne=tsnetrain[,1:2]
xtest_tsne=tsnetest[,1:2]
ytrain_tsne=tsnetrain[,3]
ytest_tsne=tsnetest[,3]

```

## logistic regression
```{r}
lr_tsne<-multinom(group~PC1+PC2,data=tsnetrain,maxit=500) 
summary(lr_tsne) 
```

## Result
```{r}
prob_tsnetrain<-predict(lr_tsne,type="probs")   
head(prob_tsnetrain)

pred_tsnetrain<-predict(lr_tsne,type="class")  
head(pred_tsnetrain)
table<-table(predicted=pred_tsnetrain,Actual=tsnetrain$group) 
table
Accuracy<-sum(diag(table))/sum(table) 
Accuracy
```

## Save
```{r}
v1=PC[1:452,1]
v2=PC[1:452,2]
d2=matrix(rep(0,452*452),nrow = 452)
for (i in 1:452){
  for (j in 1:452){
    d2[i,j]=(v1[i]-v1[j])^2+(v2[i]-v2[j])^2
  }
}

write.csv(d2,"tsne3.csv")
```

### rPCA

## create data frame with 0 rows and 3 columns
```{r}
PC <- data.frame(matrix(ncol = 3, nrow = 0))
```

## Reduction
```{r}
for (i in 1:10){

  index<-which(y$group==target_gp[i])

  new=rate[index,]
  new=t(new)

  X_rpca=robpca(new,k=2)
  X_rpca = X_rpca$loadings
  X_rpca=as.data.frame(X_rpca)

  PC<-rbind(PC,data.frame("x"=X_rpca$PC1, "y"=X_rpca$PC2, "group"<-rep(target_gp[i],gp[i])))

}
```

## provide column names
```{r}
colnames(PC)<-c("PC1","PC2","group")
```

## Plot
```{r}
ggplot(data = PC,aes(x=PC1, y = PC2, col=factor(group)))+geom_point()
```

## splitting
```{r}
set.seed(42)
sample=sample(c(TRUE, FALSE), nrow(PC), replace=TRUE, prob=c(0.7,0.3))
isotrain  <- PC[sample, ]
isotest   <- PC[!sample, ]
xtrain_iso=isotrain[,1:2]
xtest_iso=isotest[,1:2]
ytrain_iso=isotrain[,3]
ytest_iso=isotest[,3]
```

#logistic regression
```{r}
lr_iso<-multinom(group~PC1+PC2,data=isotrain,maxit=500) #maxit:iteration times
summary(lr_iso) 
```

## Error and accuracy
```{r}
prob_isotrain<-predict(lr_iso,type="probs") 
head(prob_isotrain)

pred_isotrain<-predict(lr_iso,type="class") 
head(pred_isotrain)

#Confusion Matrix
table<-table(predicted=pred_isotrain,Actual=isotrain$group) 
table

#Accuracy
Accuracy<-sum(diag(table))/sum(table)  
Accuracy

#lr_iso=glm(group~PC1+PC2,data=isotrain)

#summary(lr_iso)

```

## Save data to csv
```{r}
v1=PC[1:452,1]
v2=PC[1:452,2]
d2=matrix(rep(0,452*452),nrow = 452)
for (i in 1:452){
  for (j in 1:452){
    d2[i,j]=(v1[i]-v2[j])^2
  }
}

write.csv(d2,"LLE2.csv")
```

### sPCA

#create data frame with 0 rows and 3 columns
```{r}
PC <- data.frame(matrix(ncol = 3, nrow = 0))
```

## Reduction
```{r}
for (i in 1:10){

  index<-which(y$group==target_gp[i])

  new=rate[index,]
  new=t(new)

  X_spca=spca(new,k=2)
  X_spca = X_spca$loadings
  X_spca=as.data.frame(X_spca)

  PC<-rbind(PC,data.frame("x"=X_spca$V1, "y"=X_spca$V2, "group"<-rep(i,gp[i])))

}
```

## provide column names
```{r}
colnames(PC)<-c("PC1","PC2","group")
```

## Plot
```{r}
ggplot(data = PC,aes(x=PC1, y = PC2, col=factor(group)))+geom_point()
```

## splitting
```{r}
set.seed(42)
sample=sample(c(TRUE, FALSE), nrow(PC), replace=TRUE, prob=c(0.7,0.3))
spcatrain  <- PC[sample, ]
spcatest   <- PC[!sample, ]
xtrain_spca=spcatrain[,1:2]
xtest_spca=spcatest[,1:2]
ytrain_spca=spcatrain[,3]
ytest_spca=spcatest[,3]
```

## regression
```{r}
lr_spca<-multinom(group~PC1+PC2,data=spcatrain,maxit=500) 
summary(lr_spca) 
```

## Result
```{r}
prob_spcatrain<-predict(lr_spca,type="probs")  
head(prob_spcatrain)

pred_spcatrain<-predict(lr_spca,type="class")  
head(pred_spcatrain)
table<-table(predicted=pred_spcatrain,Actual=spcatrain$group) 
table
Accuracy<-sum(diag(table))/sum(table)
Accuracy

```

## Save
```{r}
v1=PC[1:452,1]
v2=PC[1:452,2]
d2=matrix(rep(0,452*452),nrow = 452)
for (i in 1:452){
  for (j in 1:452){
    d2[i,j]=(v1[i]-v1[j])^2+(v2[i]-v2[j])^2
  }
}

write.csv(d2,"spca3.csv")
```

## Plot in groups
```{r}
#demo<-data.frame("x"=X_spca$V1, "y"=X_spca$V2, "group"<-c(rep(1,6),rep(2,20),rep(3,11)))
#colnames(demo)<-c("x","y","z")
#ggplot(data = demo,aes(x=x, y = y, col=factor(z)))+geom_point()

```