### 2023-04-16 course/csic5011/2023/project1/Hao-Hu-Li-Liu/Code2.Rmd

---
title: "project1"
author: "Yifan HAO"
date: "2023-04-04"
output: html_document
---

data loading
===

```{r}
library(R.matlab)
datmat<-readMat("C:/Users/Lenovo/Desktop/project1/snp452-data.mat")

z<-datmat$X
x<-datmat$X[-1,]
for(i in 1:452){
  x[,i]<-diff(z[,i])/z[-1258,i]
}
x<-as.data.frame(x)
y<-as.data.frame(t(array(unlist(datmat$stock),dim = c(3,452))))#name/full name/group
colnames(y)<-c("name","detail","group")
colnames(x)<-y[,1]

```

```{r}
#bar plot in group
library(ggplot2)
z<-unique(y$group)
n<-length(z)
gp<-rep(0,n)
for(i in 1:n){
  gp[i]<-length(which(y$group==z[i]))
}

count_bar<-data.frame("group"=z,"frequency"=gp)
ggplot(data = count_bar,aes(x=factor(group),y=frequency,fill=factor(group)))+geom_bar(stat = "identity")+xlab("Groups")+ylab("Frequency")+theme(axis.text.x = element_blank(),legend.key.size=unit(10,"pt"))+guides(fill=guide_legend(title = "Group"))

```

data explore and target select
===

```{r}
target<-rep(0,10)
for(i in 1:10){
  index<-count_bar$group[i]
  print(index)
  print(which(y$group==index))
  target[i]<-which(y$group==index)[1]
}
```

```{r}
#ts plot
for(i in 1:10){
  plot(1:1257,as.numeric(x[,target[i]]),type = "l",main = count_bar$group[i])
}
```

```{r}
#heatmap in-group
library(ggcorrplot)
target<-target[-9]
target_gp<-count_bar$group[-9]
h<-matrix(rep(0,452*452),ncol = 452)
for(i in 1:452){
  for(j in i:452){
    h[i,j]<-cor(x[,i],x[,j])
    h[j,i]<-h[i,j]
  }
}

for(i in 1:9){
  index<-which(y$group==target_gp[i])
  plt<-ggcorrplot(h[index,index])
  print(plt)
}
```

function define
===

```{r}
library(forecast)

#pca method
sele.pca<-function(x,xt,nf=0){#input the predictors(n*t) and the number of factors(option)
  n<-nrow(x)
  t<-ncol(x)
  tt<-ncol(xt)
  dm<-apply(x,1,mean)
  xd<-x-matrix(rep(dm,t),ncol = t)#n*t
  ss<-1/n*t(xd)%*%xd
  de<-eigen(ss)
  te<-de$values[1]*min(n,t)^(-1/3)
  num<-length(which(de$values>=te))
  if(nf!=0){
    num<-nf
  }
  fac<-sqrt(t)*matrix(as.numeric(de$vectors[,1:num]),nrow = t)#factor in sample t*nf
  rr<-solve(t(fac)%*%fac,tol=.Machine$double.eps^3)%*%t(fac)%*%t(xd)#loading nf*n
  fac<-t(fac)
  if(tt>1){
    dtm<-apply(xt,1,mean)
    xtd<-xt-matrix(rep(dtm,tt),ncol = tt)
  }else{
    xtd<-as.matrix(xt,ncol=1)-matrix(dm,ncol = 1)
  }
  facf<-solve(rr%*%t(rr),tol=.Machine$double.eps^3)%*%rr%*%xtd#factor out of sample
  l<-list("num"=num,"x.in"=fac,"x.out"=facf)
  return(l)
}

#scaled-pca method
sele.spca<-function(y,x,xt,nf=0){#input the predictors(n*t) and the number of factors(option)
  n<-nrow(x)
  t<-ncol(x)
  tt<-ncol(xt)
  z<-scale(t(x),scale = F)#center
  gam<-rep(0,n)
  #cat("a")
  for(i in 1:n){
    if(sum(abs(as.numeric(z[,i])))>0){
      gam[i]<-lm(y~as.numeric(z[,i]))$coef[2]
      #gam[i]<-qc.calcu(as.numeric(z[,i]),y,tau = tau)
    }
    #cat(i,gam[i],"\n")
  }
  xh<-diag(gam)%*%x#weighted
  xth<-diag(gam)%*%xt
  dm<-apply(xh,1,mean)
  xd<-xh-matrix(rep(dm,t),ncol = t)#n*t
  ss<-1/n*t(xd)%*%xd
  de<-eigen(ss)
  te<-de$values[1]*min(n,t)^(-1/3)
  num<-length(which(de$values>=te))
  if(nf!=0){
    num<-nf
  }
  fac<-sqrt(t)*matrix(as.numeric(de$vectors[,1:num]),nrow = t)#factor in sample
  rr<-solve(t(fac)%*%fac,tol=.Machine$double.eps^3)%*%t(fac)%*%t(xd)#loading
  fac<-t(fac)
  if(tt>1){
    dtm<-apply(xth,1,mean)
    xtd<-xth-matrix(rep(dtm,tt),ncol = tt)
  }else{
    xtd<-as.matrix(xth,ncol=1)-matrix(dm,ncol = 1)
  }
  facf<-solve(rr%*%t(rr),tol=.Machine$double.eps^3)%*%rr%*%xtd#factor out of sample
  l<-list("num"=num,"x.in"=fac,"x.out"=facf)
  return(l)
}

#target-pca method
sele.tpca<-function(y,x,xt,nf=0,dis){#input the predictors(n*t) and the number of factors(option)
  n<-nrow(x)
  t<-ncol(x)
  tt<-ncol(xt)
  nn<-ceiling(t/log(t))
  z<-scale(t(x),scale = F)#center
  gam<-rep(1,n)
  #gr<-rep(0,n)
  #for(i in 1:n){
    #if(sum(abs(as.numeric(z[,i])))>0){
      #gam[i]<-lm(y~as.numeric(z[,i]))$coef[2]^2
    #}
    #cat(i,gam[i],"\n")
  #}
  gr<-rank(dis)
  #gr<-n+1-rank(gam)
  #gam[which(rank(gr)>=40)]<-rep(0,length(which(rank(gr)>=40)))
  xh<-x[which(gr<=nn),]#weighted
  xth<-xt[which(gr<=nn),]
  #xh<-diag(gam)%*%x#weighted
  #xth<-diag(gam)%*%xt
  dm<-apply(xh,1,mean)
  xd<-xh-matrix(rep(dm,t),ncol = t)#n1*t
  ss<-1/nn*t(xd)%*%xd
  de<-eigen(ss)
  te<-de$values[1]*min(n,t)^(-1/3)
  num<-length(which(de$values>=te))
  if(nf!=0){
    num<-nf
  }
  fac<-sqrt(t)*matrix(as.numeric(de$vectors[,1:num]),nrow = t)#factor in sample
  rr<-solve(t(fac)%*%fac,tol=.Machine$double.eps^3)%*%t(fac)%*%t(xd)#loading
  fac<-t(fac)
  if(tt>1){
    dtm<-apply(xth,1,mean)
    xtd<-xth-matrix(rep(dtm,tt),ncol = tt)
  }else{
    xtd<-as.matrix(xth,ncol=1)-matrix(dm,ncol = 1)
  }
  facf<-solve(rr%*%t(rr),tol=.Machine$double.eps^3)%*%rr%*%xtd#factor out of sample
  l<-list("num"=num,"x.in"=fac,"x.out"=facf)
  return(l)
}

```

forecasting
===

```{r}
frst<-function(x,y,pt,fn,bk,dis){ 
#(x,y) is data; pt is the forecasting point; fn is factor number max value; bk is the lag value; dis is the feature distance in different alg 
  t<-length(y)
  lt<-matrix(rep(0,7*(t-pt+1)),ncol = 7)#baseline/pca/spca/tpca-mds/tpca-iso/tpca-tsne/tpca-spca
  ll<-list(lt)
  for(k in 2:fn){
    ll[[k]]<-lt
  }
  if(bk==0){
    bk<-1
    y_bk<-matrix(rep(0,bk*length(y[(bk+1):t])),ncol = bk)
  }else{
    y_bk<-matrix(rep(0,bk*length(y[(bk+1):t])),ncol = bk)
    for(i in 1:bk){
      y_bk[,i]<-y[(bk+1-i):(t-i)]
    }
  }
  for(i in pt:t){
    #print(paste(i,"-th"))
    st<-bk+1+i-pt
    ed<-i-1
    #or
    if(bk>0){
      fit.or<-lm(y[st:ed]~y_bk[(st-bk):(ed-bk),])
      cc<-fit.or$coefficients
      cc[which(is.na(cc))]<-0
      for(bn in 1:fn){
        ll[[bn]][i-pt+1,1]<-sum(c(1,as.matrix(y_bk[i-bk,]))*cc)
      }
    }

    #print("baseline")

    #pca
    para.pca<-sele.pca(t(as.matrix(x[(st-1):(ed-1),])),as.matrix(x[i-1,]),nf=fn)
    for(bn in 1:fn){
      if(bn==1){
        fit.pca<-lm(y[st:ed]~para.pca$x.in[1,]+y_bk[(st-bk):(ed-bk),])
      }else{
        fit.pca<-lm(y[st:ed]~t(para.pca$x.in[1:bn,])+y_bk[(st-bk):(ed-bk),])
      }
      cc<-fit.pca$coefficients
      cc[which(is.na(cc))]<-0
      ll[[bn]][i-pt+1,2]<-sum(c(1,as.numeric(para.pca$x.out[1:bn,]),as.numeric(y_bk[i-bk,]))*cc)
    }
    #print("pca")

    #spca
    para.spca<-sele.spca(y[st:ed],t(as.matrix(x[(st-1):(ed-1),])),as.matrix(x[i-1,]),nf=fn)
    for(bn in 1:fn){
      if(bn==1){
        fit.spca<-lm(y[st:ed]~para.spca$x.in[1,]+y_bk[(st-bk):(ed-bk),])
      }else{
        fit.spca<-lm(y[st:ed]~t(para.spca$x.in[1:bn,])+y_bk[(st-bk):(ed-bk),])
      }
      cc<-fit.spca$coefficients
      cc[which(is.na(cc))]<-0
      ll[[bn]][i-pt+1,3]<-sum(c(1,as.numeric(para.spca$x.out[1:bn,]),as.numeric(y_bk[i-bk,]))*cc)
    }
    #print("spca")

    #tpca-mds
    para.tpca<-sele.tpca(y[st:ed],t(as.matrix(x[(st-1):(ed-1),])),as.matrix(x[i-1,]),nf=fn,dis = dis[,1])
    for(bn in 1:fn){
      if(bn==1){
        fit.tpca<-lm(y[st:ed]~para.tpca$x.in[1,]+y_bk[(st-bk):(ed-bk),])
      }else{
        fit.tpca<-lm(y[st:ed]~t(para.tpca$x.in[1:bn,])+y_bk[(st-bk):(ed-bk),])
      }
      cc<-fit.tpca$coefficients
      cc[which(is.na(cc))]<-0
      ll[[bn]][i-pt+1,4]<-sum(c(1,as.numeric(para.tpca$x.out[1:bn,]),as.numeric(y_bk[i-bk,]))*cc)
    }
    #print("tpca")

    #tpca-iso
    para.tpca<-sele.tpca(y[st:ed],t(as.matrix(x[(st-1):(ed-1),])),as.matrix(x[i-1,]),nf=fn,dis = dis[,2])
    for(bn in 1:fn){
      if(bn==1){
        fit.tpca<-lm(y[st:ed]~para.tpca$x.in[1,]+y_bk[(st-bk):(ed-bk),])
      }else{
        fit.tpca<-lm(y[st:ed]~t(para.tpca$x.in[1:bn,])+y_bk[(st-bk):(ed-bk),])
      }
      cc<-fit.tpca$coefficients
      cc[which(is.na(cc))]<-0
      #print("cc-iso")
      ll[[bn]][i-pt+1,5]<-sum(c(1,as.numeric(para.tpca$x.out[1:bn,]),as.numeric(y_bk[i-bk,]))*cc)
    }
    #print("tpca")

    #tpca-tsne
    para.tpca<-sele.tpca(y[st:ed],t(as.matrix(x[(st-1):(ed-1),])),as.matrix(x[i-1,]),nf=fn,dis = dis[,3])
    for(bn in 1:fn){
      if(bn==1){
        fit.tpca<-lm(y[st:ed]~para.tpca$x.in[1,]+y_bk[(st-bk):(ed-bk),])
      }else{
        fit.tpca<-lm(y[st:ed]~t(para.tpca$x.in[1:bn,])+y_bk[(st-bk):(ed-bk),])
      }
      cc<-fit.tpca$coefficients
      cc[which(is.na(cc))]<-0
      #print("cc-tsne")
      ll[[bn]][i-pt+1,6]<-sum(c(1,as.numeric(para.tpca$x.out[1:bn,]),as.numeric(y_bk[i-bk,]))*cc)
    }
    #print("tpca")

    #tpca-spca
    para.tpca<-sele.tpca(y[st:ed],t(as.matrix(x[(st-1):(ed-1),])),as.matrix(x[i-1,]),nf=fn,dis = dis[,4])
    for(bn in 1:fn){
      if(bn==1){
        fit.tpca<-lm(y[st:ed]~para.tpca$x.in[1,]+y_bk[(st-bk):(ed-bk),])
      }else{
        fit.tpca<-lm(y[st:ed]~t(para.tpca$x.in[1:bn,])+y_bk[(st-bk):(ed-bk),])
      }
      cc<-fit.tpca$coefficients
      cc[which(is.na(cc))]<-0
      #print("cc-spca")
      ll[[bn]][i-pt+1,7]<-sum(c(1,as.numeric(para.tpca$x.out[1:bn,]),as.numeric(y_bk[i-bk,]))*cc)
    }
    #print("tpca")

  }
  return(ll)
}
```

```{r}
for(i in 1:9){
  aa<-auto.arima(x[,target[i]],max.p = 20,max.q = 20,max.d = 20,ic="bic")
  print(i)
  print(aa)
}
#get the result that lag==1 is reasonable
```

```{r}
tsne<-read.csv("tsne3.csv")[,-1]
spca<-read.csv("spca3.csv")[,-1]
mds<-read.csv("mds3.csv")[,-1]
iso<-read.csv("iso3.csv")[,-1]
```

```{r}
#total 1257; start from 755(pt == 755, bk == 1
re<-data.frame(matrix(rep(0,7*7*9),ncol = 7))
colnames(re)<-c("baseline","pca","spca","tpca-mds","tpca-iso","tpca-tsne","tpca-spca")
re$fac<-rep(1:7,9)
re$index<-rep(target_gp,each=7)

for(i in 1:9){
  print(i)
  dis<-matrix(c(as.numeric(mds[target[i],-target[i]]),as.numeric(iso[target[i],-target[i]]),as.numeric(tsne[target[i],-target[i]]),as.numeric(spca[target[i],-target[i]])),ncol = 4)
  ll<-frst(as.matrix(x[,-target[i]]),as.numeric(x[,target[i]]),pt=755,fn=7,bk=1,dis = dis)
  for(j in 1:7){
    for(k in 1:7){
      re[(i-1)*7+j,k]<-mean((ll[[j]][,k] - x[755:1257,target[i]])^2)
    }
  }
}
```

```{r}
write.csv(re, file = "mse_result_0.2")
```