### 2023-04-18 course/csic5011/2023/project1/Yan-Yan-Lai/Untitled2.Rmd
project1

---
title: "Untitled"
output: html_document
date: "2023-04-01"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
crime_data <- read_csv("crime2.csv") 
crime_data
```

```{r}
crime_data %>% 
  mutate(date = if_else(date_wa == ".", date_my, date_wa)) %>% 
  filter(date != ".") %>% 
  mutate(police_number = sworn+civil) %>% 
  select(elecyear:web, -mayor, -termlim, -web, -date_wa, -date_my, -sworn, -civil, -jid, date, police_number) %>% 
  drop_na() %>% 
  mutate(crime_rate = (murder+rape+robbery+assault+burglary+larceny+auto)/citypop) %>% 
  select(-murder, -rape, -robbery, -assault, -burglary, -larceny, -auto) %>% 
  view()
```

```{r}
crime_data_mod <- crime_data %>% 
  mutate(police_number = sworn+civil) %>% 
  select(name:web, -city, -statenam, -state, -censdist, -mayor, -termlim, -web, -date_wa, -date_my, -sworn, -civil, -jid, police_number) %>% 
  drop_na() %>% 
  mutate(crime_rate = (murder+rape+robbery+assault+burglary+larceny+auto)/citypop) %>% 
  select(-murder, -rape, -robbery, -assault, -burglary, -larceny, -auto)

crime_data_mod

```

```{r}
library(tidymodels)

pca_rec <- recipe(crime_rate ~ ., data = crime_data_mod) %>%
  update_role(name, year, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_pca(all_predictors(), num_comp = 5)

pca_prep <- prep(pca_rec)

pca_prep
```

```{r}
sdev <- pca_prep$steps[[2]]$res$sdev

sdev/sum(sdev)
```

```{r}
pca_prep$steps
```

```{r}
tidied_pca <- tidy(pca_prep, 2)

tidied_pca %>%
  filter(component %in% paste0("PC", 1:5)) %>%
  mutate(component = fct_inorder(component)) %>%
  ggplot(aes(value, terms, fill = terms)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~component, nrow = 1) +
  labs(y = NULL)

ggsave("1.png")
```

```{r}
library(tidytext)

tidied_pca %>%
  filter(component %in% paste0("PC", 1:4)) %>%
  group_by(component) %>%
  top_n(8, abs(value)) %>%
  ungroup() %>%
  mutate(terms = reorder_within(terms, abs(value), component)) %>%
  ggplot(aes(abs(value), terms, fill = value > 0)) +
  geom_col() +
  facet_wrap(~component, scales = "free_y") +
  scale_y_reordered() +
  labs(
    x = "Absolute value of contribution",
    y = NULL, fill = "Positive?"
  )

ggsave("2.png")
```

```{r}
juice(pca_prep)
```

```{r}
juice(pca_prep) %>%
  ggplot(aes(PC01, PC02, label = str_c(name, year, sep = ","))) +
  geom_point(aes(color = crime_rate), alpha = 1, size = 2) +
  geom_text(check_overlap = FALSE, vjust = -0.8, hjust = "outward", size = 2.3) +
  labs(color = NULL) +
  scale_color_gradient(
    low = "dark blue",
    high = "tomato",
  ) +
  theme(legend.key.height = unit(1.5, "cm"))

ggsave("8.png", scale = 6)
```

```{r}
juice(pca_prep) %>%
  filter(year == 1975) %>% 
  ggplot(aes(PC01, PC02, label = name)) +
  geom_point(aes(color = crime_rate), alpha = 1, size = 2) +
  geom_text(check_overlap = TRUE, vjust = -1, size = 2.3) +
  labs(color = NULL) +
  scale_color_gradient(
    low = "dark blue",
    high = "tomato",
  ) +
  theme(legend.key.height = unit(1.5, "cm"))

ggsave("3.png")
```

```{r}
juice(pca_prep) %>%
  filter(year == 1985) %>% 
  ggplot(aes(PC01, PC02, label = name)) +
  geom_point(aes(color = crime_rate), alpha = 1, size = 2) +
  geom_text(check_overlap = TRUE, vjust = -1, size = 2.3) +
  labs(color = NULL) +
  scale_color_gradient(
    low = "dark blue",
    high = "tomato",
  ) +
  theme(legend.key.height = unit(1.5, "cm"))

ggsave("6.png")
```

```{r}
juice(pca_prep) %>%
  filter(name == "atlanta") %>% 
  ggplot(aes(PC01, PC02, label = year)) +
  geom_point(aes(color = crime_rate), alpha = 1, size = 2) +
  geom_text(check_overlap = TRUE, vjust = -1.1, size = 2) +
  labs(color = NULL) +
  scale_color_gradient(
    low = "dark blue",
    high = "tomato",
  ) +
  theme(legend.key.height = unit(1.5, "cm"))

ggsave("4.png")
```

```{r}
juice(pca_prep) %>%
  filter(name == "chicago") %>% 
  ggplot(aes(PC01, PC02, label = year)) +
  geom_point(aes(color = crime_rate), alpha = 1, size = 2) +
  geom_text(check_overlap = TRUE, vjust = -1.1, size = 2) +
  labs(color = NULL) +
  scale_color_gradient(
    low = "dark blue",
    high = "tomato",
  ) +
  theme(legend.key.height = unit(1.5, "cm"))

ggsave("5.png")
```

```{r}
isomap_rec <- recipe(crime_rate ~ ., data = crime_data_mod) %>%
  update_role(name, year, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_isomap(all_predictors())

isomap_prep <- prep(isomap_rec)

isomap_prep
```

```{r}
juice(isomap_prep)
```

```{r}
juice(isomap_prep) %>%
  ggplot(aes(Isomap1, Isomap2, label = str_c(name, year, sep = ","))) +
  geom_point(aes(color = crime_rate), alpha = 1, size = 2) +
  geom_text(check_overlap = FALSE, vjust = -0.8, hjust = "outward", size = 2.3) +
  labs(color = NULL) +
  scale_color_gradient(
    low = "dark blue",
    high = "tomato",
  ) +
  theme(legend.key.height = unit(1.5, "cm"))

ggsave("7.png", scale = 6)
```

```{r}
library(embed)

umap_rec <- recipe(crime_rate ~ ., data = crime_data_mod) %>%
  update_role(name, year, new_role = "id") %>%
  step_normalize(all_predictors()) %>%
  step_umap(all_predictors(), num_comp = 5)

umap_prep <- prep(umap_rec)

umap_prep
```

```{r}
juice(umap_prep)
```

```{r}
juice(umap_prep) %>%
  ggplot(aes(UMAP1, UMAP2, label = str_c(name, year, sep = ","))) +
  geom_point(aes(color = crime_rate), alpha = 1, size = 2) +
  geom_text(check_overlap = FALSE, vjust = -0.8, hjust = "outward", size = 2.3) +
  labs(color = NULL) +
  scale_color_gradient(
    low = "dark blue",
    high = "tomato",
  ) +
  theme(legend.key.height = unit(1.5, "cm"))

ggsave("11.png", scale = 6)
```

```{r}
library(ggforce)

juice(pca_prep) %>%
  select(crime_rate:PC05) %>% 
  ggplot() +
  geom_autopoint(aes(color = crime_rate), alpha = 0.4, size = 0.5) +
  geom_autodensity(alpha = .3) +
  facet_matrix(vars(-crime_rate), layer.diag = 2) +
  labs(color = NULL) +
  scale_color_gradient(
    low = "dark blue",
    high = "tomato",
  ) +
  theme(legend.key.height = unit(1.5, "cm"))

ggsave("9.png")
```

```{r}
juice(isomap_prep) %>%
  select(crime_rate:Isomap5) %>% 
  ggplot() +
  geom_autopoint(aes(color = crime_rate), alpha = 0.4, size = 0.5) +
  geom_autodensity(alpha = .3) +
  facet_matrix(vars(-crime_rate), layer.diag = 2) +
  labs(color = NULL) +
  scale_color_gradient(
    low = "dark blue",
    high = "tomato",
  ) +
  theme(legend.key.height = unit(1.5, "cm"))

ggsave("10.png")
```

```{r}
juice(umap_prep) %>%
  select(crime_rate:UMAP5) %>% 
  ggplot() +
  geom_autopoint(aes(color = crime_rate), alpha = 0.4, size = 0.5) +
  geom_autodensity(alpha = .3) +
  facet_matrix(vars(-crime_rate), layer.diag = 2) +
  labs(color = NULL) +
  scale_color_gradient(
    low = "dark blue",
    high = "tomato",
  ) +
  theme(legend.key.height = unit(1.5, "cm"))

ggsave("12.png")
```

```{r}
lm_spec <- 
  linear_reg() %>% 
  set_engine("lm")

lm_wf <- workflow(pca_rec, lm_spec)

lm_fit <- lm_wf %>% fit(crime_data_mod)

lm_fit
```

```{r}
lm_fit$fit$fit$fit %>% summary()
```

```{r}
null_rec <- recipe(crime_rate ~ ., data = crime_data_mod) %>%
  update_role(name, year, new_role = "id")

lm_wf2 <- workflow(null_rec, lm_spec)

lm_fit2 <- lm_wf2 %>% fit(crime_data_mod)

lm_fit2
```

```{r}
lm_fit2$fit$fit$fit %>% summary()
```

```{r}
library(Rdimtools)

mds <- do.mds(crime_data_mod %>% 
  select(-name, -year, -crime_rate) |> 
  as.matrix() , ndim = 5)

mds$Y %>% 
  as_tibble() %>% 
  rename(MDS1 = V1, MDS2 = V2,  MDS3 = V3,  MDS4 = V4,  MDS5 = V5)

```