### 2023-05-23 course/csic5011/2023/project1/1.rebuttal/SHEN_rebuttal.md
csic5011

# Rebuttal

Thanks for your reading and all of your review comments! Here are my feedbacks.

> CUI Yifan:
>
> 1. K-means method is used without theoretical background description or proper citation and
> 2. the number of pages exceeds the maximum requirement.

Regarding the use of the k-means method, I apologize for not providing a more detailed theoretical background description or proper citation. I will revise the manuscript to provide a more thorough explanation of the method and appropriate citations.

Regarding the length of the paper, I acknowledge that it exceeds the maximum requirement. However, I believe that the additional information and analysis provided in the paper are necessary for a comprehensive understanding of my findings. I will work to condense the paper while still ensuring that all relevant information is included. Thanks for your feedback.

> Gerry DUNDA:
>
> - The paper does not discuss the limitations of the statistical methods used for the analysis, such as PCA, MDS, and t-SNE. While these methods can be powerful tools for exploring high-dimensional data, they have certain assumptions and limitations that may affect their results.

I agree that every statistical method has its own assumptions and limitations, including PCA, MDS, and t-SNE. In my paper, I did mention some limitations of the data and methods used, such as the small sample size and potential biases in the data collection. However, I acknowledge that more discussion on the limitations of the statistical methods could have been included. Regarding the limitations of PCA, I am aware that PCA assumes a linear relationship between variables, and it may not be the most suitable method for non-linear data. Also, the interpretation of the principal components may not always be straightforward, and there could be noise in the data that affects the results. Similarly, MDS and t-SNE have their own assumptions and limitations, such as the choice of distance metrics and the effects of noise and outliers. I could have discussed these limitations in more detail, and I appreciate your feedback on this. Thanks for your feedback and I will keep this in mind for future research.

> HUANG Zhanmiao:
>
> - The contents and explored questions are relatively basic and simple. The investigation can go deeper.

While I agree that further investigations can always be conducted, I believe that my study provides valuable insights into the migration patterns and genetic diversity of human populations. I think my work can serve as a foundation for future studies to build upon and deepen my understanding of these complex topics.

> LI Yakun:
>
> 1. It would be better if potential limitations of these methods are discussed in the report.
> 2. However, there are some minor typos that could be corrected.

Regarding question 1, please check my feedbacks for Gerry. Thanks.

Regarding question 2, I appreciate your feedback and will carefully review and correct any errors that I find. Thank you for bringing this to my attention.

> LIU Yonglin:
>
> - The report does not discuss potential limitations of the PCA method or the assumptions made during the analysis.

Please check my feedbacks for Gerry. Thanks.

> LUO Yuanhui
>
> - How human migration history is revealed by the relationship between SNPs and geographical information can be clearer.

Regarding the clarity of how human migration history is revealed by the relationship between SNPs and geographical information, I agree that this could be further clarified. In particular, I could provide more detail on the specific SNPs that were found to be informative for inferring population history, and how they relate to known migration events in human history. Additionally, I could provide more detail on the statistical methods used to infer these relationships and how they contribute to our understanding of human migration history. Thanks for your valuable feedback.

> TENG Fei
>
> - It seems that the sample size will affect the experiment result. But there exists few experiments to test relative large sample size.

I appreciate your comment and agree that sample size can have an impact on the results of experiments. In this study, we acknowledge the limitation of our relatively small sample size of 1043 individuals. However, we believe that our results provide valuable insights into human migration history based on the available data. As for experiments with larger sample sizes, I hope that future studies can build upon my work and explore this topic further. Thanks.

> ZHONG Guangzheng:
>
> - The figures should follows the text that they are metioned.

Thanks for your feedback. I have reviewed all figures and their captions, no such issue has been found yet. Please elaborate more if you found any. Thanks.