### 2023-05-23 course/csic5011/2023/project2/Li-Liu-Hao-Hu/Reduction.Rmd

---
title: "Reduction, visualization and classification"
author: "Liu Yonglin"
date: "2023-04-09"
output: html_document
---

```{r setup, include=FALSE}
library(R.matlab)
library(RDRToolbox)
library(ggplot2)
library(ggcorrplot)
library(Rtsne)
#library(rpca)
library(rospca)
library(sparsepca)
#library(lle)
library(randomForest)
library(pROC)
library(nnet)
```

## set seed
```{r}
set.seed(42)
```

## Load data
```{r}
library(fbi)
#md <- fredmd(file = "monthly.csv", transform = TRUE)
qd <- fredqd(file = "quarterly.csv",transform = TRUE)
#View(md)
View(qd)
qd_dis<-fredqd_description
#md_dis<-fredmd_description

for(i in 2:247){
  aa<-which(is.na(qd[61:252,i]))
  if(length(aa)>0){
    print("r.v")
    print(i)
    print(aa+60)
  }
}

#remove 22,76,92,135,138,142,168,179,180,182,188,232
#from 1974Q1 to 2021Q4
#total: 192 variables number: 235-1=234
qdt<-qd[61:252,-c(22,76,92,135,138,142,168,179,180,182,188,232)]

#将qdt中缺失值替换为0
for(i in 2:234){
  qdt[which(is.na(qdt[,i])),i]<-0
}

am<-as.matrix(qdt[,-1])

path<-getwd()
path<-paste(path,"index-group.csv",sep = "/")
ig <- read.csv(path, header=TRUE)
```

```{r}
col_names <- colnames(am[,2:10])
years<- qdt$date

y=years
for (i in 1:8){
  y=c(y,years)
}

test <- qdt[1:192, 2:10]
abc = unlist(test)
abc=as.data.frame(abc)
xxx=abc[1:1728,]

index <- rep(col_names, each = 192)

bb<- data.frame(years = y, value = xxx, index = index)
ggplot(data = bb, aes(x=years,  y=value, colour=index)) +geom_line() 
```

all <- data.frame(matrix(ncol = 2, nrow = 0))

### ISOMAP

##Create data frame with 0 rows and 2 columns
```{r}
PC <- data.frame(matrix(ncol = 2, nrow = 0))
```

## Data reduction and creating new dataframe PC with 4 principle components
```{r}
X_ISO<-Isomap(data=am, dims=2, k=4)
X_ISO<-as.data.frame(X_ISO)

PC<-rbind(PC,data.frame("x"=X_ISO$dim2.1, "y"=X_ISO$dim2.2))

```

## provide column names
```{r}
colnames(PC)<-c("PC1","PC2")
```

all <- rbind(all, PC)

## splitting
```{r}
set.seed(42)
sample=sample(c(TRUE, FALSE), nrow(PC), replace=TRUE, prob=c(0.7,0.3))
isotrain  <- PC[sample, ]
isotest   <- PC[!sample, ]
xtrain_iso=isotrain[,1:2]
xtest_iso=isotest[,1:2]
ytrain_iso=isotrain[,3]
ytest_iso=isotest[,3]
```

#logistic regression
```{r}
lr_iso<-multinom(group~PC1+PC2,data=isotrain,maxit=500) #maxit:iteration times
summary(lr_iso) 
```

## Error and accuracy
```{r}
prob_isotrain<-predict(lr_iso,type="probs") 
head(prob_isotrain)

pred_isotrain<-predict(lr_iso,type="class") 
head(pred_isotrain)

#Confusion Matrix
table<-table(predicted=pred_isotrain,Actual=isotrain$group) 
table

#Accuracy
Accuracy<-sum(diag(table))/sum(table)  
Accuracy

#lr_iso=glm(group~PC1+PC2,data=isotrain)

#summary(lr_iso)

```

## Plot
```{r}

ggplot(data = PC,aes(  x= PC1, y = PC2) )+ geom_point(color="red",size=2) + coord_cartesian(xlim=c(-1e+06,1e+06),ylim=c(-250000,250000))+ labs(title = "ISOMAP") + theme(plot.title = element_text(hjust = 0.5))
```

## Plot2
```{r}
rr=ig$group
PC$"group"<-rr[1:192]
ggplot(data = PC,aes(  x= PC1, y = PC2,col=group) ) + geom_point() + coord_cartesian(xlim=c(-1e+06,1e+06),ylim=c(-250000,250000))+ labs(title = "ISOMAP") + theme(plot.title = element_text(hjust = 0.5))
```

## Save
```{r}
write.csv(PC,"isomap-2pc.csv")

v1=PC[1:192,1]
v2=PC[1:192,2]
d2=matrix(rep(0,192*192),nrow = 192)
for (i in 1:192){
  for (j in 1:192){
    d2[i,j]=(v1[i]-v1[j])^2+(v2[i]-v2[j])^2
  }
}

write.csv(d2,"isomap-dist-2pc.csv")
```

##Create data frame with 0 rows and 4 columns
```{r}
PC <- data.frame(matrix(ncol = 4, nrow = 0))
```

## Data reduction and creating new dataframe PC with 4 principle components
```{r}
X_ISO<-Isomap(data=am, dims=4, k=4)
X_ISO<-as.data.frame(X_ISO)

PC<-rbind(PC,data.frame("1"=X_ISO$dim4.1, "2"=X_ISO$dim4.2, "3"=X_ISO$dim4.3, "4"=X_ISO$dim4.4))

```

## provide column names
```{r}
colnames(PC)<-c("PC1","PC2","PC3","PC4")
```

## Save-pc4
```{r}
write.csv(PC,"isomap-4pc.csv")

d2 <- dist(PC)

d2 <- as.data.frame(as.matrix(d2))
write.csv(d2,"isomap-dist-4pc.csv")
```

##Create data frame with 0 rows and 6 columns
```{r}
PC <- data.frame(matrix(ncol = 6, nrow = 0))
```

## Data reduction and creating new dataframe PC with 6 principle components
```{r}
X_ISO<-Isomap(data=am, dims=6, k=4)
X_ISO<-as.data.frame(X_ISO)

PC<-rbind(PC,data.frame("1"=X_ISO$dim6.1, "2"=X_ISO$dim6.2, "3"=X_ISO$dim6.3, "4"=X_ISO$dim6.4, X_ISO$dim6.5, X_ISO$dim6.6))

```

## provide column names
```{r}
colnames(PC)<-c("PC1","PC2","PC3","PC4","PC5","PC6")
```

## Save-pc4
```{r}
write.csv(PC,"isomap-6pc.csv")

d2 <- dist(PC)

d2 <- as.data.frame(as.matrix(d2))
write.csv(d2,"isomap-dist-6pc.csv")
```

### LLE

##Create data frame with 0 rows and 2 columns
```{r}
PC <- data.frame(matrix(ncol = 2, nrow = 0))
```

## Data reduction
```{r}
X_LLE<-LLE(data=am, dim=2, k=2)
X_LLE<-as.data.frame(X_LLE)

PC<-rbind(PC,data.frame("x"=X_LLE$V1, "y"=X_LLE$V2))
```

##provide column names
```{r}
colnames(PC)<-c("PC1","PC2")
```

all <- rbind(all, PC)

##Plot
```{r}
ggplot(data = PC,aes(x=PC1, y = PC2, col=factor(group)))+geom_point(color="darkorange",size=2.5)+ coord_cartesian(xlim=c(-2,2),ylim=c(-1.7,3))+ labs(title = "LLE") + theme(plot.title = element_text(hjust = 0.5))
```
## Plot2
```{r}
rr=ig$group
PC$"group"<-rr[1:192]
ggplot(data = PC,aes(  x= PC1, y = PC2,col=group) ) + geom_point() + coord_cartesian(xlim=c(-2,2),ylim=c(-1.7,3))+ labs(title = "LLE") + theme(plot.title = element_text(hjust = 0.5))
```

## splitting
```{r}
set.seed(42)
sample=sample(c(TRUE, FALSE), nrow(PC), replace=TRUE, prob=c(0.7,0.3))
isotrain  <- PC[sample, ]
isotest   <- PC[!sample, ]
xtrain_iso=isotrain[,1:2]
xtest_iso=isotest[,1:2]
ytrain_iso=isotrain[,3]
ytest_iso=isotest[,3]
```

#logistic regression
```{r}
lr_iso<-multinom(group~PC1+PC2,data=isotrain,maxit=500) #maxit:iteration times
summary(lr_iso) 
```

## Error and accuracy
```{r}
prob_isotrain<-predict(lr_iso,type="probs") 
head(prob_isotrain)

pred_isotrain<-predict(lr_iso,type="class") 
head(pred_isotrain)

#Confusion Matrix
table<-table(predicted=pred_isotrain,Actual=isotrain$group) 
table

#Accuracy
Accuracy<-sum(diag(table))/sum(table)  
Accuracy

#lr_iso=glm(group~PC1+PC2,data=isotrain)

#summary(lr_iso)

```

## Save data to csv
```{r}
write.csv(PC,"lle-2pc.csv")

v1=PC[1:192,1]
v2=PC[1:192,2]
d2=matrix(rep(0,192*192),nrow = 192)
for (i in 1:192){
  for (j in 1:192){
    d2[i,j]=(v1[i]-v1[j])^2+(v2[i]-v2[j])^2
  }
}

write.csv(d2,"lle-dist-2pc.csv")
```

##Create data frame with 0 rows and 4 columns
```{r}
PC <- data.frame(matrix(ncol = 4, nrow = 0))
```

## Data reduction
```{r}
X_LLE<-LLE(data=am, dim=4, k=2)
X_LLE<-as.data.frame(X_LLE)

PC<-rbind(PC,data.frame("x"=X_LLE$V1, "y"=X_LLE$V2, X_LLE$V3, X_LLE$V4))
```

##provide column names
```{r}
colnames(PC)<-c("PC1","PC2","PC3","PC4")
```

## Save data to csv
```{r}
write.csv(PC,"lle-4pc.csv")

d2 <- dist(PC)

d2 <- as.data.frame(as.matrix(d2))

write.csv(d2,"lle-dist-4pc.csv")
```

##Create data frame with 0 rows and 6 columns
```{r}
PC <- data.frame(matrix(ncol = 6, nrow = 0))
```

## Data reduction
```{r}
X_LLE<-LLE(data=am, dim=6, k=2)
X_LLE<-as.data.frame(X_LLE)

PC<-rbind(PC,data.frame("x"=X_LLE$V1, "y"=X_LLE$V2, X_LLE$V3, X_LLE$V4, X_LLE$V5, X_LLE$V6))
```

##provide column names
```{r}
colnames(PC)<-c("PC1","PC2","PC3","PC4","PC5","PC6")
```

## Save data to csv
```{r}
write.csv(PC,"lle-6pc.csv")

d2 <- dist(PC)

d2 <- as.data.frame(as.matrix(d2))

write.csv(d2,"lle-dist-6pc.csv")
```

### MDS

## create data frame with 0 rows and 2 columns
```{r}
PC <- data.frame(matrix(ncol = 2, nrow = 0))
```

## Data reduction
```{r}

dist = dist(am,p = 2)
X_MDS= cmdscale(dist)
X_MDS<-as.data.frame(X_MDS)

PC<-rbind(PC,data.frame(X_MDS$V1, X_MDS$V2))

```

## provide column names
```{r}
colnames(PC)<-c("PC1","PC2")
```

all <- rbind(all, PC)

## Plot
```{r}
ggplot(data = PC,aes(x=PC1, y = PC2, col=factor(group)))+geom_point(color="darkgreen",size=2.5)+ coord_cartesian(xlim=c(-1e+06,1e+06),ylim=c(-2e+05,1e+05))+ labs(title = "MDS") + theme(plot.title = element_text(hjust = 0.5))
```

## Plot2
```{r}
rr=ig$group
PC$"group"<-rr[1:192]
ggplot(data = PC,aes(  x= PC1, y = PC2,col=group) ) + geom_point()+ coord_cartesian(xlim=c(-1e+06,1e+06),ylim=c(-2e+05,1e+05))+ labs(title = "MDS") + theme(plot.title = element_text(hjust = 0.5))
```

## splitting
```{r}
set.seed(42)
sample=sample(c(TRUE, FALSE), nrow(PC), replace=TRUE, prob=c(0.7,0.3))
mdstrain  <- PC[sample, ]
mdstest   <- PC[!sample, ]
xtrain_mds=mdstrain[,1:2]
xtest_mds=mdstest[,1:2]
ytrain_mds=mdstrain[,3]
ytest_mds=mdstest[,3]
```

## logistic regression
```{r}
lr_mds<-multinom(group~PC1+PC2,data=mdstrain,maxit=500) #maxit表示迭代次数
summary(lr_mds) 
```

## Accuracy
```{r}
prob_mdstrain<-predict(lr_mds,type="probs")   
head(prob_mdstrain)

pred_mdstrain<-predict(lr_mds,type="class")   
head(pred_mdstrain)
table<-table(predicted=pred_mdstrain,Actual=mdstrain$group)
table
Accuracy<-sum(diag(table))/sum(table)
Accuracy

```

## Save
```{r}
write.csv(PC,"mds-2pc.csv")

v1=PC[1:192,1]
v2=PC[1:192,2]
d2=matrix(rep(0,192*192),nrow = 192)
for (i in 1:192){
  for (j in 1:192){
    d2[i,j]=(v1[i]-v1[j])^2+(v2[i]-v2[j])^2
  }
}

write.csv(d2,"mds-dist-2pc.csv")
```

## create data frame with 0 rows and 4 columns
```{r}
PC <- data.frame(matrix(ncol = 4, nrow = 0))
```

## Data reduction
```{r}

dist = dist(am,p = 2)
X_MDS= cmdscale(dist,k=4)
X_MDS<-as.data.frame(X_MDS)

PC<-rbind(PC,data.frame(X_MDS$V1, X_MDS$V2, X_MDS$V3, X_MDS$V4))

```

##provide column names
```{r}
colnames(PC)<-c("PC1","PC2","PC3","PC4")
```
## Save
```{r}
write.csv(PC,"mds-4pc.csv")

d2 <- dist(PC)

d2 <- as.data.frame(as.matrix(d2))

write.csv(d2,"mds-dist-4pc.csv")

```

## create data frame with 0 rows and 6 columns
```{r}
PC <- data.frame(matrix(ncol = 6, nrow = 0))
```

## Data reduction
```{r}

dist = dist(am,p = 2)
X_MDS= cmdscale(dist,k=6)
X_MDS<-as.data.frame(X_MDS)

PC<-rbind(PC,data.frame(X_MDS$V1, X_MDS$V2, X_MDS$V3, X_MDS$V4, X_MDS$V5, X_MDS$V6))

```

##provide column names
```{r}
colnames(PC)<-c("PC1","PC2","PC3","PC4","PC5","PC6")
```
## Save
```{r}
write.csv(PC,"mds-6pc.csv")

d2 <- dist(PC)

d2 <- as.data.frame(as.matrix(d2))

write.csv(d2,"mds-dist-6pc.csv")

```

### TSNE

## create data frame with 0 rows and 2 columns
```{r}
PC <- data.frame(matrix(ncol = 2, nrow = 0))
```

#reduction
```{r}

tSNE_out = Rtsne(am,check_duplicates = FALSE, perplexity = 1)
X_tSNE = tSNE_out$Y
X_tSNE<-as.data.frame(X_tSNE)

PC<-rbind(PC,data.frame("x"=X_tSNE$V1, "y"=X_tSNE$V2))

```

##provide column names
```{r}
colnames(PC)<-c("PC1","PC2")
```

all <- rbind(all, PC)

##Plot
```{r}
ggplot(data = PC,aes(x=PC1, y = PC2, col=factor(group)))+geom_point(colour="blue",size=2.2)+ labs(title = "TSNE") + theme(plot.title = element_text(hjust = 0.5))
```

## Plot2
```{r}
rr=ig$group
PC$"group"<-rr[1:192]
ggplot(data = PC,aes(  x= PC1, y = PC2,col=group) ) + geom_point()+ labs(title = "TSNE") + theme(plot.title = element_text(hjust = 0.5))
```

## splitting
```{r}
set.seed(42)
sample=sample(c(TRUE, FALSE), nrow(PC), replace=TRUE, prob=c(0.7,0.3))
tsnetrain  <- PC[sample, ]
tsnetest   <- PC[!sample, ]
xtrain_tsne=tsnetrain[,1:2]
xtest_tsne=tsnetest[,1:2]
ytrain_tsne=tsnetrain[,3]
ytest_tsne=tsnetest[,3]

```

## logistic regression
```{r}
lr_tsne<-multinom(group~PC1+PC2,data=tsnetrain,maxit=500) 
summary(lr_tsne) 
```

## Result
```{r}
prob_tsnetrain<-predict(lr_tsne,type="probs")   
head(prob_tsnetrain)

pred_tsnetrain<-predict(lr_tsne,type="class")  
head(pred_tsnetrain)
table<-table(predicted=pred_tsnetrain,Actual=tsnetrain$group) 
table
Accuracy<-sum(diag(table))/sum(table) 
Accuracy
```

## Save
```{r}
write.csv(PC,"tsne-2pc.csv")

v1=PC[1:192,1]
v2=PC[1:192,2]
d2=matrix(rep(0,192*192),nrow = 192)
for (i in 1:192){
  for (j in 1:192){
    d2[i,j]=(v1[i]-v1[j])^2+(v2[i]-v2[j])^2
  }
}

write.csv(d2,"tsne-dist-2pc.csv")
```

## create data frame with 0 rows and 4 columns
```{r}
PC <- data.frame(matrix(ncol = 4, nrow = 0))
```

#reduction(max_dim=3)
```{r}

tSNE_out = Rtsne(am, dim = 3, check_duplicates = FALSE, perplexity = 1)
X_tSNE = tSNE_out$Y
X_tSNE<-as.data.frame(X_tSNE)

PC<-rbind(PC,data.frame("x"=X_tSNE$V1, "y"=X_tSNE$V2))

```

### rPCA

## create data frame with 0 rows and 2 columns
```{r}
PC <- data.frame(matrix(ncol = 2, nrow = 0))
```

## Reduction
```{r}

am1=t(am)

X_rpca=robpca(am1,k=2)
X_rpca = X_rpca$loadings
X_rpca=as.data.frame(X_rpca)

PC<-rbind(PC,data.frame("x"=X_rpca$PC1, "y"=X_rpca$PC2))

```

## provide column names
```{r}
colnames(PC)<-c("PC1","PC2")
```

all <- rbind(all, PC)

## Plot
```{r}
ggplot(data = PC,aes(x=PC1, y = PC2, col=factor(group)))+geom_point(colour="purple",size=2.5)+ coord_cartesian(xlim=c(-0.2,0.3),ylim=c(-0.3,0.25))+ labs(title = "rPCA") + theme(plot.title = element_text(hjust = 0.5))
```

## Plot2
```{r}
rr=ig$group
PC$"group"<-rr[1:192]
ggplot(data = PC,aes(  x= PC1, y = PC2,col=group) ) + geom_point()+ coord_cartesian(xlim=c(-0.2,0.3),ylim=c(-0.3,0.25))+ labs(title = "rPCA") + theme(plot.title = element_text(hjust = 0.5))
```

## splitting
```{r}
set.seed(42)
sample=sample(c(TRUE, FALSE), nrow(PC), replace=TRUE, prob=c(0.7,0.3))
isotrain  <- PC[sample, ]
isotest   <- PC[!sample, ]
xtrain_iso=isotrain[,1:2]
xtest_iso=isotest[,1:2]
ytrain_iso=isotrain[,3]
ytest_iso=isotest[,3]
```

#logistic regression
```{r}
lr_iso<-multinom(group~PC1+PC2,data=isotrain,maxit=500) #maxit:iteration times
summary(lr_iso) 
```

## Error and accuracy
```{r}
prob_isotrain<-predict(lr_iso,type="probs") 
head(prob_isotrain)

pred_isotrain<-predict(lr_iso,type="class") 
head(pred_isotrain)

#Confusion Matrix
table<-table(predicted=pred_isotrain,Actual=isotrain$group) 
table

#Accuracy
Accuracy<-sum(diag(table))/sum(table)  
Accuracy

#lr_iso=glm(group~PC1+PC2,data=isotrain)

#summary(lr_iso)

```

## Save data to csv
```{r}
write.csv(PC,"rpca-2pc.csv")

v1=PC[1:192,1]
v2=PC[1:192,2]
d2=matrix(rep(0,192*192),nrow = 192)
for (i in 1:192){
  for (j in 1:192){
    d2[i,j]=(v1[i]-v1[j])^2+(v2[i]-v2[j])^2
  }
}

write.csv(d2,"rpca-dist-2pc.csv")
```

## create data frame with 0 rows and 4 columns
```{r}
PC <- data.frame(matrix(ncol = 4, nrow = 0))
```

## Reduction
```{r}

am1=t(am)

X_rpca=robpca(am1,k=4)
X_rpca = X_rpca$loadings
X_rpca=as.data.frame(X_rpca)

PC<-rbind(PC,data.frame(X_rpca$PC1, X_rpca$PC2, X_rpca$PC3, X_rpca$PC4))

```

## provide column names
```{r}
colnames(PC)<-c("PC1","PC2","PC3","PC4")
```

## Save
```{r}
write.csv(PC,"rpca-4pc.csv")

d2 <- dist(PC)

d2 <- as.data.frame(as.matrix(d2))

write.csv(d2,"rpca-dist-4pc.csv")

```

## create data frame with 0 rows and 6 columns
```{r}
PC <- data.frame(matrix(ncol = 6, nrow = 0))
```

## Reduction
```{r}

am1=t(am)

X_rpca=robpca(am1,k=6)
X_rpca = X_rpca$loadings
X_rpca=as.data.frame(X_rpca)

PC<-rbind(PC,data.frame(X_rpca$PC1, X_rpca$PC2, X_rpca$PC3, X_rpca$PC4, X_rpca$PC5, X_rpca$PC6))

```

## provide column names
```{r}
colnames(PC)<-c("PC1","PC2","PC3","PC4","PC5","PC6")
```

## Save
```{r}
write.csv(PC,"rpca-6pc.csv")

d2 <- dist(PC)

d2 <- as.data.frame(as.matrix(d2))

write.csv(d2,"rpca-dist-6pc.csv")

```
### sPCA

#create data frame with 0 rows and 2 columns
```{r}
PC <- data.frame(matrix(ncol = 2, nrow = 0))
```

## Reduction
```{r}

am2=t(am)

X_spca=spca(am2,k=2)
X_spca = X_spca$loadings
X_spca=as.data.frame(X_spca)

PC<-rbind(PC,data.frame("x"=X_spca$V1, "y"=X_spca$V2))

```

## provide column names
```{r}
colnames(PC)<-c("PC1","PC2")
```

all <- rbind(all, PC)

## Plot
```{r}
ggplot(data = PC,aes(x=PC1, y = PC2, col=factor(group)))+geom_point(colour="#FF6699",size=2.5)+ coord_cartesian(xlim=c(0,0.17),ylim=c(-0.3,0.15))+ labs(title = "sPCA") + theme(plot.title = element_text(hjust = 0.5))
```
## Plot2
```{r}
rr=ig$group
PC$"group"<-rr[1:192]
ggplot(data = PC,aes(  x= PC1, y = PC2,col=group) ) + geom_point()+ coord_cartesian(xlim=c(0,0.17),ylim=c(-0.3,0.15))+ labs(title = "sPCA") + theme(plot.title = element_text(hjust = 0.5))
```
## splitting
```{r}
set.seed(42)
sample=sample(c(TRUE, FALSE), nrow(PC), replace=TRUE, prob=c(0.7,0.3))
spcatrain  <- PC[sample, ]
spcatest   <- PC[!sample, ]
xtrain_spca=spcatrain[,1:2]
xtest_spca=spcatest[,1:2]
ytrain_spca=spcatrain[,3]
ytest_spca=spcatest[,3]
```

## regression
```{r}
lr_spca<-multinom(group~PC1+PC2,data=spcatrain,maxit=500) 
summary(lr_spca) 
```

## Result
```{r}
prob_spcatrain<-predict(lr_spca,type="probs")  
head(prob_spcatrain)

pred_spcatrain<-predict(lr_spca,type="class")  
head(pred_spcatrain)
table<-table(predicted=pred_spcatrain,Actual=spcatrain$group) 
table
Accuracy<-sum(diag(table))/sum(table)
Accuracy

```

## Save
```{r}
write.csv(PC,"spca-2pc.csv")

v1=PC[1:192,1]
v2=PC[1:192,2]
d2=matrix(rep(0,192*192),nrow = 192)
for (i in 1:192){
  for (j in 1:192){
    d2[i,j]=(v1[i]-v1[j])^2+(v2[i]-v2[j])^2
  }
}

write.csv(d2,"spca-dist-2pc.csv")
```

## create data frame with 0 rows and 4 columns
```{r}
PC <- data.frame(matrix(ncol = 4, nrow = 0))
```

## Reduction
```{r}

am1=t(am)

X_spca=spca(am1,k=4)
X_spca = X_spca$loadings
X_spca=as.data.frame(X_spca)

PC<-rbind(PC,data.frame(X_spca$PC1, X_spca$PC2, X_spca$PC3, X_spca$PC4))

```

## provide column names
```{r}
colnames(PC)<-c("PC1","PC2","PC3","PC4")
```

## Save
```{r}
write.csv(PC,"spca-4pc.csv")

d2 <- dist(PC)

d2 <- as.data.frame(as.matrix(d2))

write.csv(d2,"spca-dist-4pc.csv")

```

## create data frame with 0 rows and 6 columns
```{r}
PC <- data.frame(matrix(ncol = 6, nrow = 0))
```

## Reduction
```{r}

am1=t(am)

X_spca=spca(am1,k=6)
X_spca = X_spca$loadings
X_spca=as.data.frame(X_spca)

PC<-rbind(PC,data.frame(X_spca$PC1, X_spca$PC2, X_spca$PC3, X_spca$PC4, X_spca$PC5, X_spca$PC6))

```

## provide column names
```{r}
colnames(PC)<-c("PC1","PC2","PC3","PC4","PC5","PC6")
```

## Save
```{r}
write.csv(PC,"spca-6pc.csv")

d2 <- dist(PC)

d2 <- as.data.frame(as.matrix(d2))

write.csv(d2,"spca-dist-6pc.csv")
```

## Plot in groups
```{r}
index <- rep(c("ISOMAP","LLE", "MDS", "TSNE", "RPCA", "SPCA"), each = 192)
all$method <- index
```

```{r}
ggplot(data = all,aes(x=PC1, y = PC2, col=index))+geom_point(size=0.5)
```

#Plot

library(ggplot2)
library(scales)

path<-getwd()
path<-paste(path,"final_mse.csv",sep = "/")
df <- read.csv(path, header=TRUE)

df1 <- df[df$index == "GDPC1", ]
df11 <- df1[, 2:10]
df11 <- t(df11)

rows <- c(1, 2, 5, 8, 9)
new_df <- t(df11[rows, ])

mse2 = unlist(df1)

mse2=as.data.frame(mse2)

factor <- rep(1:9, each = 7)
mse3 <- mse2[8:70, ]
index <- rep(c("baseline","pca", "spca", "tpca-mds", "tpca-iso", "tpca-tsne", "tpca-spca", "tpca-rpca", "tpca-lle"), each = 7)

q=c(1,2,3,4,5,6,7)

t=q
for (i in 1:8){
  t=c(t,q)
}

aa <- data.frame(factor = t, mse = mse3, index = index)

aa$mse <- format(aa[,2], scientific = TRUE)

ggplot(data = aa, aes(x=factor,y=mse,colour=index)  ) + geom_line()

df2 <- df[df$index == "INDPRO", ]
df22 <- t(df2[, 2:10])

mse2 = unlist(df2)

mse2=as.data.frame(mse2)

factor <- rep(1:9, each = 7)
mse <- mse2[8:70, ]
index <- rep(c("baseline","pca", "spca", "tpca-mds", "tpca-iso", "tpca-tsne", "tpca-spca", "tpca-rpca", "tpca-lle"), each = 7)

q=c(1,2,3,4,5,6,7)

t=q
for (i in 1:8){
  t=c(t,q)
}

aa <- data.frame(factor = t, mse = mse, index = index)

ggplot(data = aa, aes(x=factor,  y=mse , colour=index)) +geom_point()

df2 <- df[df$index == "HWIURATIOx", ]
df22 <- t(df2[, 2:10])

mse2 = unlist(df2)

mse2=as.data.frame(mse2)

factor <- rep(1:9, each = 7)
mse3 <- mse2[8:70, ]
index <- rep(c("baseline","pca", "spca", "tpca-mds", "tpca-iso", "tpca-tsne", "tpca-spca", "tpca-rpca", "tpca-lle"), each = 7)

q=c(1,2,3,4,5,6,7)

t=q
for (i in 1:8){
  t=c(t,q)
}

aa <- data.frame(factor = t, mse = mse3, index = index)

ggplot(data = aa, aes(x=factor,  y=mse , colour=index)) +geom_point()

df2 <- df[df$index == "DFSARG3Q086SBEA", ]
df22 <- t(df2[, 2:10])

mse2 = unlist(df2)

mse2=as.data.frame(mse2)

factor <- rep(1:9, each = 7)
mse <- mse2[8:70, ]
index <- rep(c("baseline","pca", "spca", "tpca-mds", "tpca-iso", "tpca-tsne", "tpca-spca", "tpca-rpca", "tpca-lle"), each = 7)

q=c(1,2,3,4,5,6,7)

t=q
for (i in 1:8){
  t=c(t,q)
}

aa <- data.frame(factor = t, mse = mse, index = index)

ggplot(data = aa, aes(x=factor,  y=mse , colour=index)) +geom_point()